### 2024-06-25 src/tex/alg2.tex
build: organize directory structure and refine environment

\caption{Generating new candidate rules}\label{alg:generate_cands}
\begin{algorithmic}[1]
  \Function{GenerateCands}{$\mathcal{A},x$}
  \IIf{$\mathcal{A}=\emptyset$}{\Return{$\{\mathit{true}\}$}}
    \Comment{An initial empty rule always returns $\mathit{true}$}
  \State$\cands\gets\emptyset$
  \ForAll{$A\in\mathcal{A}$}
  \ForAll{$a\in (T(x)\setminus A)$}
  \State$\cands\gets\bar{\mathcal{A}}\cup(A\wedge a)$
  \Comment{Get a new rule by adding a new predicate $a$ to $A$}
  \EndFor%
  \EndFor%
  \State\Return{$\cands$}
  \EndFunction%
\end{algorithmic}